# Coursera Data Science Certificates 
## Reproducible Research Course Project 1  

Student Name: Min Wang

Date Created: 07 August 2015  

### Load libraries
```{r}
library(knitr)
library(data.table)
library(plyr)
library(ggplot2)
library(zoo)
library(scales)
library(Hmisc)
```
  

### Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
```
  

### Questions 1  
#### What is mean total number of steps taken per day? (Ignore the missing values)  
##### (1) Calculate the total number of steps taken per day
```{r}
answer_1_1 <- ddply(data, .(date), summarise, total_step_daily=mean(steps, na.rm=TRUE))
answer_1_1
```

##### (2) Make a histogram of the total number of steps taken each day.
```{r}
hist(answer_1_1$total_step_daily, main = paste("Histogram of" , "the total number of steps taken each day", sep=" "), xlab="the total number of steps taken each day")
```

##### (3) Calculate and report the mean and median of the total number of steps taken per day
```{r}
answer_1_3 <- summary(answer_1_1$total_step_daily)
answer_1_3
```

### Questions 2
#### What is the average daily activity pattern?
##### (1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
answer_2_1 <- ddply(data, .(interval), summarise, average_step_interval=mean(steps, na.rm=TRUE))
answer_2_1
ggplot(answer_2_1, aes(x=interval, y=average_step_interval)) + geom_line() + ggtitle("Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days") + xlab("5-minute interval") + ylab("average number of steps taken")
```
##### (2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
answer_2_2 <- answer_2_1[which.max(answer_2_1$average_step_interval), ]
answer_2_2
```

### Questions 3
#### Imputing missing values
##### (1) Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
answer_3_1 <- length(is.na(data$steps))
answer_3_1
```
##### (2) Devise a strategy for filling in all of the missing values in the dataset.
Imputation Strategy: Use mean steps of the day.
```{r}
is.na(answer_2_1$average_step_interval)
tmp1 <- as.data.table(data)
tmp2 <- as.data.table(answer_2_1)
setkey(tmp1, interval)
setkey(tmp2, interval)
answer_3_2 <- merge(tmp1, tmp2, all.x=TRUE)
answer_3_2
```
##### (3) Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
tmp <- as.data.frame(answer_3_2)
answer_3_3 <- tmp[, c(4, 3, 1)]
colnames(answer_3_3) <- colnames(data)
answer_3_3
```
##### (4) Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
par(mfrow=c(1, 2), mar=c(2, 4, 2, 1), oma=c(2, 0, 2, 0))
subtitle1 <- paste("Median:", answer_1_3[3], "; Mean:", answer_1_3[4], sep="")
hist(answer_1_1$total_step_daily, main=subtitle1)
answer_3_4 <- ddply(answer_3_3, .(date), summarise, total_step_daily=mean(steps, na.rm=TRUE))
tmp <- summary(answer_3_4$total_step_daily)
subtitle2 <- paste("Median:", tmp[3], "; Mean:", tmp[4], sep="")
hist(answer_3_4$total_step_daily, main=subtitle2)
mtext(paste("Histogram of" , "the total number of steps taken each day", sep=" "),
      outer = TRUE, cex=1.5)
```

### Questions 4
#### Are there differences in activity patterns between weekdays and weekends?
##### (1) Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
answer_4_1 <- answer_3_3
answer_4_1$date <- as.Date(answer_4_1$date, "%Y-%m-%d")
answer_4_1 <- mutate(answer_4_1, date_of_week=weekdays(answer_4_1$date))
answer_4_1 <- mutate(answer_4_1, weekday=as.factor(ifelse(weekdays(answer_4_1$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")))
```
##### (2) Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
answer_4_2 <- ddply(answer_4_1, .(interval, weekday), summarise, average_step_interval=mean(steps, na.rm=TRUE))
ggplot(answer_4_2, aes(x=interval, y=average_step_interval)) + facet_grid(weekday~.) + geom_line(colour="sky blue") + xlab("Interval") + ylab("Number of steps taken")
```

