---
title: "The fatal and economic impacts of tornadoes"
---

## Synopsis
Use U.S. National Oceanic and Atmospheric Administration's (NOAA) storm data, we found tornadoes not only caused highest number of fatality (5633) and injuries (91346), but also resulted in 3312276.7 billion dollars economic loss.

## Load Libraries
```{r}
library(knitr)
library(data.table)
library(plyr)
library(ggplot2)
library(zoo)
library(scales)
library(Hmisc)
library(reshape2)
```

## Read Data
```{r}
data <- read.csv("storm_data.csv")
```

## Data Processing
### Questions 1
#### Across the United States, which types of events are most harmful with respect to population health?
```{r}
answer_1 <- ddply(data, .(EVTYPE), summarise, total_fatality=sum(FATALITIES, na.rm=TRUE), total_injuries=sum(INJURIES, na.rm=TRUE))
table_1  <- answer_1[with(answer_1, order(total_fatality, total_injuries, decreasing=TRUE)), ]
```

### Question 2
#### Across the United States, which types of events have the greatest economic consequences?
```{r}
answer_2 <- ddply(data, .(EVTYPE), summarise, total=sum(PROPDMG+CROPDMG, na.rm=TRUE))
table_2  <- answer_2[with(answer_2, order(total, decreasing=TRUE)), ]
```

## Results
### Human casualties
From table below, we can see tornadoes (TORNADO) caused the largest number of fatality and injuries in total respectively.
```{r}
head(table_1)
```

### Economic impact
From figure below, we can see tornadoes (TORNADO) caused the largest economic loss.
```{r}
ggplot(table_2[1:10, ], aes(x=EVTYPE, y=total)) + geom_bar(stat="identity") + labs(title = "Economic Loss") + labs(x = "Event") + labs(y = "Billion Dollars") + theme(plot.title = element_text(lineheight=.8, face="bold"), axis.text.x=element_text(angle=45,vjust=1,hjust=1))
```
